---
title: "Xaringan Notes"
author: "Rhian Davies"
date: "25 October 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My personal notes for creating xaringan slides. 

## Working offline

To make slides work offline, you need to download a copy of remark.js in advance, because xaringan uses the online version by default. You can use `xaringan::summon_remark()` to download the latest or a specified version of remark.js. By default, it is downloaded to libs/remark-latest.min.js.

Then change the chakra option in the YAML metadata to point to this file, e.g.,
```{yaml chakra}
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
```

## Exporting to PDF

The `webshot()` function can convert using a link to a local html. To do this, you'll use the `file` protocol. Here's a simple way to point to your local file: 

```{r}
# install.packages("webshot")
library(webshot)
file_name <- paste0("file://", normalizePath("my_xaringan.html"))
```

You can then: 

```{r, eval = FALSE} 
webshot(file_name, "mypdf.pdf")
```

